Kindly Bent to Free Us

### 2019-07-10 kindly/lang.tex
Fix region annotation and make sure borrow capture is allowed.

Environments are also used to encode several linearity-related properties
through \emph{environment-wide constraints}.
Here we present how to control both unused variables and
closure captures through such environment constraints, noted $\Cleq{\E}{k}$.
The intuitive meaning of $\Cleq{\E}{k}$ is that for all bindings
$\bvar{x}{\tau}$ in $\E$,
if type $\tau$ has kind $k'$ (noted $\inferSK{C}{\E}{\tau}{k'}$)
then the following constraints should hold: $\Cleq{k'}{k}$.
While this definition is incomplete, it is sufficient to understand
its use in the {\sc Abs} and {\sc Var} rules shown in \cref{selectrules:binders}.
%
The {\sc Var} rule looks up the type scheme of the variable $x$ in
the environment $\E$
and instantiate a concrete type through the substitution $\unif$. It then
checks that the other bindings in $\E$ can safely be discarded by
ensuring they are not linear. This is achieved through the constraint
$\Cleq{\E\Sdel{x}}{\kaff_\infty}$ which replaces the traditional
``weakening'' rule found in many type systems.

The {\sc Abs} rule must ensures that if the body of the lambda captures
linear or affine variables, the content will be reflected in the kind
annotated on the arrow, such as $\tau_2 \tarr{k} \tau_1$.
In order to do so, we constraint all the bindings of $\E$ to be less than $k$.
If one of the binding of $\E$ contains a type whose kind is affine,
we will derive the constraints $\Cleq{\kaff}{k}$, which ensures that
the kind on the arrow is at least affine.
Note that $x$ is \emph{not} included in $\E$: since it's the argument of the
closure, its use should not induce a capture.

@@

\begin{figure}[!h]
  \begin{minipage}{0.4\linewidth}
    \centering
    \begin{mathpar}
      \ruleSDRegion
    \end{mathpar}
    \caption{The {\sc Region} rule}
    \label{selectrules:region}
  \end{minipage}\hfill
  \begin{minipage}{0.6\linewidth}
    \centering
    \begin{tabular}
      {@{}>{$}r<{$}@{ $\vdash_e$ }
      >{$}c<{$}@{ $\rightsquigarrow_n^{x}$ }
      >{$}l<{$}
      r}

      {(\kun_n\lk k\lk\kun_\infty)}
      &{\svar[\IBORROW]{x}{\tau}^n}
      &{\bvar{\borrow[\IBORROW]{x}}{\borrowty[\IBORROW] k{\tau}}}
      &Immut\\

      (\kaff_n\lk k\lk\kaff_\infty)
      &\svar[\MBORROW]{x}{\tau}^n
      &\bvar{\borrow[\MBORROW]{x}}{\borrowty[\MBORROW] k{\tau}}
      &Mut
    \end{tabular}
    \caption{Borrowing rules for bindings -- $\lregion{C}{x}{\E}{\E'}$}
    \label{env:rule:borrow}
  \end{minipage}
\end{figure}

>>>>

Environments are also used to encode several linearity-related properties
through \emph{environment-wide constraints}.
Here we present how to control both unused variables and
closure captures through such environment constraints, noted $\Cleq{\E}{k}$.
The intuitive meaning of $\Cleq{\E}{k}$ is that for all bindings
$\bvar{x}{\tau}$ in $\E$,
if type $\tau$ has kind $k'$ (noted $\inferSK{C}{\E}{\tau}{k'}$)
then the following constraints should hold: $\Cleq{k'}{k}$.
Borrow bindings follow the exact same rules, but suspended bindings
are forbidden in such constraints.
While this definition is incomplete, it is sufficient to understand
its use in the {\sc Abs} and {\sc Var} rules shown in \cref{selectrules:binders}.
%
The {\sc Var} rule looks up the type scheme of the variable $x$ in
the environment $\E$
and instantiate a concrete type through the substitution $\unif$. It then
checks that the other bindings in $\E$ can safely be discarded by
ensuring they are not linear. This is achieved through the constraint
$\Cleq{\E\Sdel{x}}{\kaff_\infty}$ which replaces the traditional
``weakening'' rule found in many type systems.

The {\sc Abs} rule must ensures that if the body of the lambda captures
linear or affine variables, the content will be reflected in the kind
annotated on the arrow, such as $\tau_2 \tarr{k} \tau_1$.
In order to do so, we constraint all the bindings of $\E$ to be less than $k$.
If one of the binding of $\E$ contains a type whose kind is affine,
we will derive the constraints $\Cleq{\kaff}{k}$, which ensures that
the kind on the arrow is at least affine.
Capturing a borrow is perfectly allowed here: the kind of the borrow
will be reflected on the arrow which will restrict the closure
to the region of the borrow. Capturing a \emph{suspended} binding, however,
is forbidden.
Note that $x$ is \emph{not} included in $\E$: since it's the argument of the
closure, its use should not induce a capture.

@@

\begin{figure}[!h]
  \begin{minipage}{0.4\linewidth}
    \centering
    \begin{mathpar}
      \ruleSDRegion
    \end{mathpar}
    \caption{The {\sc Region} rule}
    \label{selectrules:region}
  \end{minipage}\hfill
  \begin{minipage}{0.6\linewidth}
    \centering
    % \begin{tabular}
    %   {@{}>{$}r<{$}@{ $\vdash_e$ }
    %   >{$}c<{$}@{ $\rightsquigarrow_n^{x}$ }
    %   >{$}l<{$}
    %   r}

    %   (\kun_n\lk k\lk\kun_\infty)
    %   &{\svar[\IBORROW]{x}{\tau}^n}
    %   &{\bvar{\borrow[\IBORROW]{x}}{\borrowty[\IBORROW] k{\tau}}}
    %   &Immut\\

    %   (\kaff_n\lk k\lk\kaff_\infty)
    %   &\svar[\MBORROW]{x}{\tau}^n
    %   &\bvar{\borrow[\MBORROW]{x}}{\borrowty[\MBORROW] k{\tau}}
    %   &Mut
    % \end{tabular}
    \begin{mathpar}
      \inferrule{
        \entail{C}{(\BORROW_n\lk k\lk \BORROW_\infty)}\\
        \BORROW\in\left\{\kun;\kaff\right\}
      }{
        \lregion{C}{}
        {\svar[\BORROW]{x}{\tau}^n}
        {\bvar{\borrow[\BORROW]{x}}{\borrowty[\BORROW] k{\tau}}}
      }
    \end{mathpar}
    \caption{Borrowing rules for bindings -- $\lregion{C}{x}{\E}{\E'}$}
    \label{env:rule:borrow}
  \end{minipage}
\end{figure}
